---
title: "MindWandering_PD_hallucinations"
author: "Claire O'Callaghan"
date: "May 2019"
output: html_document
---

```{r echo=FALSE, message=FALSE, warning=FALSE} 
library("knitr")
library("pander")
library("plyr")
library("rcompanion")
library("car")
library("agricolae")
library("reshape2")
library("lme4")
library("lmerTest")
library("MHTdiscrete")
library("ggpubr")

setwd("~/Desktop/Papers_in_the_works/Mind_Wandering_PD/Analysis_September_2018/")
df <- read.csv("PD_MW_data.csv", header = TRUE, quote="\"", stringsAsFactors= TRUE, strip.white = TRUE, na.strings=c("NA", "-", "?"))
attach(df)

df2 <- df2<-subset(df, Group!="Controls")
attach(df2)
```

##Summary tables demographics & neuropsychology

```{r echo=FALSE, message=FALSE, warning=FALSE}
#requires plyr
Age<-ddply(df, .(Group), summarise, mean = mean(Age, na.rm = TRUE), sd = sd(Age, na.rm = TRUE), se = sd/sqrt(length(Age)))

Education<-ddply(df, .(Group), summarise, mean = mean(Education, na.rm = TRUE), sd = sd(Education, na.rm = TRUE), se = sd/sqrt(length(Education)))

MoCA<-ddply(df2, .(Group), summarise, mean = mean(MOCA, na.rm = TRUE), sd = sd(MOCA, na.rm = TRUE), se = sd/sqrt(length(MOCA)))

Duration<-ddply(df2, .(Group), summarise, mean = mean(Duration, na.rm = TRUE), sd = sd(Duration, na.rm = TRUE), se = sd/sqrt(length(Duration)))

DDE<-ddply(df2, .(Group), summarise, mean = mean(DDE, na.rm = TRUE), sd = sd(DDE, na.rm = TRUE), se = sd/sqrt(length(DDE)))

HY<-ddply(df2, .(Group), summarise, mean = mean(HY, na.rm = TRUE), sd = sd(HY, na.rm = TRUE), se = sd/sqrt(length(HY)))

UPDRS.III<-ddply(df2, .(Group), summarise, mean = mean(UPDRS.III, na.rm = TRUE), sd = sd(UPDRS.III, na.rm = TRUE), se = sd/sqrt(length(UPDRS.III)))

BDI<-ddply(df2, .(Group), summarise, mean = mean(BDI, na.rm = TRUE), sd = sd(BDI, na.rm = TRUE), se = sd/sqrt(length(BDI)))

TMT.BA<-ddply(df2, .(Group), summarise, mean = mean(TMT.BA, na.rm = TRUE), sd = sd(TMT.BA, na.rm = TRUE), se = sd/sqrt(length(TMT.BA)))

DSB<-ddply(df2, .(Group), summarise, mean = mean(DSB, na.rm = TRUE), sd = sd(DSB, na.rm = TRUE), se = sd/sqrt(length(DSB)))

LM_retention<-ddply(df2, .(Group), summarise, mean = mean(LM_per_Ret, na.rm = TRUE), sd = sd(LM_per_Ret, na.rm = TRUE), se = sd/sqrt(length(LM_per_Ret)))

pander(Age, caption = "Age")
pander(Education, caption= "Education")
pander(MoCA, caption= "MoCA")
pander(Duration, caption= "Duration (yrs)")
pander(DDE, caption= "Dopamine Dose Equivalence")
pander(HY, caption= "Hoehn & Yahr stage")
pander(UPDRS.III, caption= "UPDRS III")
pander(BDI, caption= "Beck Depression Inventory")
pander(TMT.BA, caption= "Trail Making B-A")
pander(DSB, caption= "Digit Span Backwards")
pander(LM_retention, caption= "Logical Memory %Retention")
```

##Statistics for demographics & neuropsychology

```{r echo=FALSE, message=FALSE, warning=FALSE}
#ANOVAs for three group comparisons
#require pander for table

Age.aov <- aov(Age ~ Group, data = df)
Education.aov <- aov(Education ~ Group, data = df)

pander(Age.aov, caption="Age anova")
pander(Education.aov, caption="Education anova")

#T-tests comparing PD+VH & PD-VH
MoCA.Ttest <- t.test(MOCA~Group, data = df2)
Duration.Ttest <- t.test(Duration~Group, data = df2)
DDE.Ttest <- t.test(DDE~Group, data = df2)
HY.Ttest <- t.test(HY~Group, data = df2)
UPDRS.III.Ttest <- t.test(UPDRS.III~Group, data = df2)
BDI.Ttest <- t.test(BDI~Group, data = df2)
TMT.BA.Ttest <- t.test(TMT.BA~Group, data = df2)
DSB.Ttest <- t.test(DSB~Group, data = df2)
LM_per_Ret.Ttest <- t.test(LM_per_Ret~Group, data = df2)

pander(MoCA.Ttest, caption="Moca T-test")
pander(Duration.Ttest, caption="Duration T-test")
pander(DDE.Ttest, caption="DDE T-test")
pander(HY.Ttest, caption="Hoehn & Yahr T-test")
pander(UPDRS.III.Ttest, caption="UPDRS.III T-test")
pander(BDI.Ttest, caption="BDI T-test")
pander(TMT.BA.Ttest, caption="Trail making B-A T-test")
pander(DSB.Ttest, caption="Digits backwards T-test")
pander(LM_per_Ret.Ttest, caption="Logical memory % retention T-test")
```

##Analaysis of Mind Wandering data
###Normality checks
```{r echo=FALSE, message=FALSE, warning=FALSE}
#Check normality at each Response level
#requires rcompanion for hist plots
shapiro.test(df$Level.1)
qqnorm(df$Level.1)
qqline(df$Level.1)
plotNormalHistogram(df$Level.1)

shapiro.test(df$Level.2)
qqnorm(df$Level.2)
qqline(df$Level.2)
plotNormalHistogram(df$Level.2)

shapiro.test(df$Level.3)
qqnorm(df$Level.3)
qqline(df$Level.3)
plotNormalHistogram(df$Level.3)

shapiro.test(df$Level.4)
qqnorm(df$Level.4)
qqline(df$Level.4)
plotNormalHistogram(df$Level.4)
```

###Homogeneity of variance checks
```{r echo=FALSE, message=FALSE, warning=FALSE}
#requires car
leveneTest(Level.1~Group, data = df)
leveneTest(Level.2~Group, data = df)
leveneTest(Level.3~Group, data = df)
leveneTest(Level.4~Group, data = df)
```

###Data transformation
```{r echo=FALSE, message=FALSE, warning=FALSE}
df$Level_1_sqrt <- sqrt(df$Level.1)
df$Level_2_sqrt <- sqrt(df$Level.2)
df$Level_3_sqrt <- sqrt(df$Level.3)
df$Level_4_sqrt <- sqrt(df$Level.4)

shapiro.test(df$Level_1_sqrt)
qqnorm(df$Level_1_sqrt)
qqline(df$Level_1_sqrt)
plotNormalHistogram(df$Level_1_sqrt)

shapiro.test(df$Level_2_sqrt)
qqnorm(df$Level_2_sqrt)
qqline(df$Level_2_sqrt)
plotNormalHistogram(df$Level_2_sqrt)

shapiro.test(df$Level_3_sqrt)
qqnorm(df$Level_3_sqrt)
qqline(df$Level_3_sqrt)
plotNormalHistogram(df$Level_3_sqrt)

shapiro.test(df$Level_4_sqrt)
qqnorm(df$Level_4_sqrt)
qqline(df$Level_4_sqrt)
plotNormalHistogram(df$Level_4_sqrt)
```

###Mind Wandering frequency
####plot

```{r echo=FALSE, message=FALSE, warning=FALSE}
#df$Group <- gsub('Control', 'Controls', df$Group)

plot1 <- ggbarplot(df, x = "Group", y = "MW.frequency", 
          color = "Group", add = c("mean_se","dotplot"), palette =c("#E7B800","#00AFBB", "#FF6A6A", "#FC4E07"), add.params = list(dotsize = 0.6), position = position_dodge(0.3)) + ylab("Mind-wandering\nFrequency (%)") + scale_x_discrete(name="")+
theme(legend.title=element_blank())+
theme(axis.title.x = element_text(size=15),
axis.title.y = element_text(size=12))+
theme(legend.text=element_text(size=15))+ ylim(0, 100) 

```

###Mind Wandering frequency
####Anova
```{r echo=FALSE, message=FALSE, warning=FALSE}
#require agricolae for FLSD test
mod1 = lm(Level_4_sqrt ~ Group, data=df)
Anova(lm(Level_4_sqrt ~ Group, data=df), type = 2)
LSD.test(mod1,"Group", console = TRUE)
```

###Overall task perfromance
####data preprocessing

```{r echo=FALSE, message=FALSE, warning=FALSE}
#Transform df into long format
###Subset & melt data to long for for repeated measures set up
#requires rehape2 
attach(df)
df_levels_melted<-data.frame(Group,ID,Level_1_sqrt,Level_2_sqrt,Level_3_sqrt,Level_4_sqrt)
df_levels_melted<-melt(df_levels_melted, id=c("Group","ID"))
df_levels_melted <- setNames(df_levels_melted, c("Group","ID","Level", "Response"))

df_levels_melted_raw<-data.frame(Group,ID,Level.1,Level.2,Level.3,Level.4)
df_levels_melted_raw<-melt(df_levels_melted_raw, id=c("Group","ID"))
df_levels_melted_raw <- setNames(df_levels_melted_raw, c("Group","ID","Level", "Response"))
```

###Overall task performance
####plot

```{r echo=FALSE, message=FALSE, warning=FALSE}

plot2 <- ggbarplot(df_levels_melted_raw, x = "Level", y = "Response", 
          color = "Group", add = c("mean_se","dotplot"), palette =c("#E7B800","#00AFBB", "#FF6A6A", "#FC4E07"), add.params = list(dotsize = 0.5), position = position_dodge(0.9), width = 0.82) + ylab("Overall Response\nFrequency (%)") +
theme(legend.title=element_blank())+
theme(axis.title.x = element_text(size=15),
axis.title.y = element_text(size=12))+
theme(legend.text=element_text(size=15))+ ylim(0, 100) +  scale_x_discrete(name="", labels=c("Level.1" = "Level 1", "Level.2" = "Level 2","Level.3" = "Level 3", "Level.4" = "Level 4"))

CombinedPlots <- ggarrange(plot1, plot2, ncol = 1, nrow = 2, common.legend = TRUE)

CombinedPlots

```

###Overall task performance
####ANOVA

```{r echo=FALSE, message=FALSE, warning=FALSE}
#requires lme4 & lmerTest
aov1 <- lmer(Response ~ Level * Group + (1 | ID), data = df_levels_melted)
anova(aov1, type=2)

#Pairwise t-test to follow up main effect of Level
pairwise.t.test(df_levels_melted$Response, df_levels_melted$Level, p.adjust.method ="none")
p<- c(1.4e-06,1.6e-12,3.4e-15,0.01504,0.00085,0.35606)
Sidak.p.adjust(p, 0.05, make.decision=TRUE)


#Separate Anovas to compute simple effects after the interaction
#Use df file & compute group differences at each level
Anova(lm(Level_1_sqrt ~ Group, data=df), type = 2)
Anova(lm(Level_2_sqrt ~ Group, data=df), type = 2)
Anova(lm(Level_3_sqrt ~ Group, data=df), type = 2)
Anova(lm(Level_4_sqrt ~ Group, data=df), type = 2)
```

###Duration effects
####assumption checks

```{r echo=FALSE, message=FALSE, warning=FALSE}
#Check normality at each Response level
#requires rcompanion for hist plots
shapiro.test(df$Short)
qqnorm(df$Short)
qqline(df$Short)
plotNormalHistogram(df$Short)

shapiro.test(df$Medium)
qqnorm(df$Medium)
qqline(df$Medium)
plotNormalHistogram(df$Medium)

shapiro.test(df$Long)
qqnorm(df$Long)
qqline(df$Long)
plotNormalHistogram(df$Long)

#sqrt

df$Short_sqrt <- sqrt(df$Short)
df$Medium_sqrt <- sqrt(df$Medium)
df$Long_sqrt <- sqrt(df$Long)
```

###Duration
####data preprocessing

```{r echo=FALSE, message=FALSE, warning=FALSE}
#Transform df into long format
###Subset & melt data to long for for repeated measures set up
#requires rehape2 
attach(df)
df_duration_melted<-data.frame(Group,ID,Short_sqrt,Medium_sqrt,Long_sqrt)
df_duration_melted<-melt(df_duration_melted, id=c("Group","ID"))
df_duration_melted <- setNames(df_duration_melted, c("Group","ID","Duration", "Response"))

df_duration_melted_raw<-data.frame(Group,ID,Short,Medium,Long)
df_duration_melted_raw<-melt(df_duration_melted_raw, id=c("Group","ID"))
df_duration_melted_raw <- setNames(df_duration_melted_raw, c("Group","ID","Duration", "Response"))
```

###Duration
####plot

```{r echo=FALSE, message=FALSE, warning=FALSE}

plot3 <- ggbarplot(df_duration_melted_raw, x = "Duration", y = "Response", 
          color = "Group", add = c("mean_se","dotplot"), palette =c("#E7B800","#00AFBB", "#FC4E07","#FC4E07"), add.params = list(dotsize = 0.5), position = position_dodge(0.9), width = 0.82) + ylab("Average trial score") +
theme(legend.title=element_blank())+
theme(axis.title.x = element_text(size=15),
axis.title.y = element_text(size=15))+
theme(legend.text=element_text(size=15))

plot3

```

###Duration
####ANOVA

```{r echo=FALSE, message=FALSE, warning=FALSE}
#requires lme4 & lmerTest
aov5 <- lmer(Response ~ Duration * Group + (1 | ID), data = df_duration_melted)
anova(aov5, type=2)

#FLSD to follow up main effect of Duration
LSD.test(lm(Response~Duration, data = df_duration_melted),"Duration", console = TRUE)

#FLSD to follow up main effect of Group
LSD.test(lm(Response~Group, data = df_duration_melted),"Group", console = TRUE)
```


